\Title

### 2012-12-04 synmut.tex
Coalescent tree + C2-V5

\section{Results}
We analyze from time series of viral nucleotide sequences from single patients,
which span several years of chronic
infection~\citep{shankarappa_consistent_1999, bunnik_autologous_2008,
liu_selection_2006}. Plotting the allele frequencies against time for all
polymorphic sites, it is evident that, although nonsynonymous changes are
widespread, synonymous ones are also present in various occasions at high
frequency (see \figurename~\ref{fig:aft}).
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{Shankarappa_allele_freqs_trajectories_syn_nonsynp8}
\caption{Allele frequency trajectories of typical patient, C3-V5, nonsynonymous
(solid) and synonymous mutations (dashed lines). Most synonymous mutations are
not fixed. Colors are set according to the position of the site along the C3-V5
region (red to blue). Data from Ref.~\cite{shankarappa_consistent_1999}.}
\label{fig:aft}
\end{center}
\end{figure}
Moreover, many mutations appear to change rapidly in frequency as a flock,
especially at the lower end of the frequency spectrum. These observations lead
to the idea that linkage might be widespread at the time scale of the typical
selective sweep, i.e. a few months~\citep{neher_recombination_2010}. It seems
therefore likely that synonymous mutations are not attaining high frequencies by
genetic drift but are rather exponentially amplified by selection on linked
nonsynonymous sites, a process known as {\it genetic
draft}~\citep{neher_genetic_2011}.

To test this hypothesis, we focus on the long-time fate of such synonymous
derived alleles {\it after} they are already abundant. According to the neutral
theory of independent sites, (a) an allele starting at frequency $\nu$ will
reach fixation, in the long-time limit, with a probability $P_f(\nu) = \nu$, and
get extinct otherwise; (b) the time required to reach either boundary, fixation
or extinction, is essentialy the inverse of the population size (in
generations), hence much longer than five years for HIV~\citep{boltz_ultrasensitive_2012}.
\begin{figure}
\begin{center}
\subfloat{\includegraphics[width=0.49\linewidth]{Bunnik2008_fixmid_syn_ShankanonShanka}}
\subfloat{\includegraphics[width=0.49\linewidth]{Shankarappa_fixmid_syn_V_regions}}
\caption{Left panel: fixation probability of derived synonymous alleles is strongly
suppressed in C3-V5 versus other parts of the {\it env} gene, and of
nonsynonymous ones.
Right panel: especially hard is fixation of new alleles in conserved regions flanking the V
loops. The black dashed line is the prediction from neutral
theory, for comparison purposes. Data from
Refs.~\cite{shankarappa_consistent_1999, bunnik_autologous_2008}.}
\label{fig:fixp}
\end{center}
\end{figure}

>>>>

\section{Results}
We analyze from time series of viral nucleotide sequences from single patients,
which span several years of chronic
infection~\citep{shankarappa_consistent_1999, bunnik_autologous_2008,
liu_selection_2006}. Plotting the allele frequencies against time for all
polymorphic sites, it is evident that, although nonsynonymous changes are
widespread, synonymous ones are also present in various occasions at high
frequency (see \figurename~\ref{fig:aft}).
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{Shankarappa_allele_freqs_trajectories_syn_nonsynp8}
\caption{Allele frequency trajectories of typical patient, C2-V5, nonsynonymous
(solid) and synonymous mutations (dashed lines). Most synonymous mutations are
not fixed. Colors are set according to the position of the site along the C2-V5
region (red to blue). Data from Ref.~\cite{shankarappa_consistent_1999}.}
\label{fig:aft}
\end{center}
\end{figure}
Moreover, many mutations appear to change rapidly in frequency as a flock,
especially at the lower end of the frequency spectrum. These observations lead
to the idea that linkage might be widespread at the time scale of the typical
selective sweep, i.e. a few months~\citep{neher_recombination_2010}. It seems
therefore likely that synonymous mutations are not attaining high frequencies by
genetic drift but are rather exponentially amplified by selection on linked
nonsynonymous sites, a process known as {\it genetic
draft}~\citep{neher_genetic_2011}.

To test this hypothesis, we focus on the long-time fate of such synonymous
derived alleles {\it after} they are already abundant. According to the neutral
theory of independent sites, (a) an allele starting at frequency $\nu$ will
reach fixation, in the long-time limit, with a probability $P_f(\nu) = \nu$, and
get extinct otherwise; (b) the time required to reach either boundary, fixation
or extinction, is essentialy the inverse of the population size (in
generations), hence much longer than five years for HIV~\citep{boltz_ultrasensitive_2012}.
\begin{figure}
\begin{center}
\subfloat{\includegraphics[width=0.49\linewidth]{Bunnik2008_fixmid_syn_ShankanonShanka}}
\subfloat{\includegraphics[width=0.49\linewidth]{Shankarappa_fixmid_syn_V_regions}}
\caption{Left panel: fixation probability of derived synonymous alleles is strongly
suppressed in C2-V5 versus other parts of the {\it env} gene, and of
nonsynonymous ones.
Right panel: especially hard is fixation of new alleles in conserved regions flanking the V
loops. The black dashed line is the prediction from neutral
theory, for comparison purposes. Data from
Refs.~\cite{shankarappa_consistent_1999, bunnik_autologous_2008}.}
\label{fig:fixp}
\end{center}
\end{figure}
